---
title: "Predicting Turbulence Simulations"
author: "Enzo Moraes Mescall, Martin Olarte, Charlotte Coudert"
date: "2022-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Report should be no more than 4 pages
```


```{r libraries, echo = FALSE, results = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(splines)
library(gam)
```

```{r data}
data.train = read.csv("data-train.csv")
data.test = read.csv("data-test.csv")
```


## Introduction


## Methodology

We wanted a method to quickly perform cross validation on a candidate model and estimate its test MSE, so we created a general function with that purpose.

```{r}
cv_mse = function(input_model, k = 5, iter = 25, data = clean) {
    sum_mse = 0
    for(i in 1:iter) {
      sum_mse = sum_mse + cv.glm(input_model, data = clean, K = k)$delta[1]
    }
    avg_mse = sum_mse/iter
    return(avg_mse)
}
```


After this, we created a function that would resemble our final product and take in a tuple of St, Re, and Fr and then return the four predicted moments. We reserved the global variable names `model_mX` for this final function.

```{r}
# temp models
predictive_model = function(test_St, test_Re, test_Fr) {
  newdata = data.frame(list(St = test_St, Re = as.factor(test_Re), Fr = as.factor(test_Fr)))
  
  m1 = predict(model_m1, newdata =  newdata)
  m2 = predict(model_m2, newdata =  newdata)
  m3 = predict(model_m3, newdata =  newdata)
  m4 = predict(model_m4, newdata =  newdata)
  
  return(c(m1, m2, m3, m4))
}
```

```{r}
summary(data.train)
```

```{r}
clean = data.train %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))
```


```{r}
glm(R_moment_1~ .
```

```{r}
ggplot(data = clean, aes(x= St, y = Re, color = Fr)) +
  geom_point() + geom_jitter()
```

```{r}
ggplot(data = clean, aes(x= St)) +
  geom_histogram()
ggplot(data = clean, aes(x= Fr)) +
  geom_bar()
ggplot(data = clean, aes(x= Re)) +
  geom_bar()
```


```{r gen-linear}
lin_model <- glm(R_moment_1 ~ Re + St + Fr, data = clean)
summary(lin_model)

par(mfrow = c(2,2))
plot(lin_model)
```
## MOMENT 2


```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_1 ~ Re + Fr + poly(St, degree), data = data.train)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```


```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_1 ~ Re + Fr + poly(St, degree), data = clean)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```

```{r}
ggplot(data = clean, aes(x= St, y = R_moment_2, color = Fr)) +
  geom_point()
ggplot(data = clean, aes(x= Fr, y = R_moment_2, color = Re)) +
  geom_point()
ggplot(data = clean, aes(x= Re, y = R_moment_2, color = St)) +
  geom_point()
```


```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_2 ~ Re + Fr + poly(St, degree), data = clean)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```

``{r}
gam.m1 <- gam(R_moment_2 ~ s(St, 5) + Re+ Fr, data = clean)
preds1 <- predict(gam.m1, newdata = clean)

gam.m2 <- gam(R_moment_2 ~ ns(St, 5) + Re+ Fr, data = clean)
gam.m3 <- gam(R_moment_2 ~ s(St, 3)  + Re+ Fr, data = clean)
anova(gam.m1, gam.m2, gam.m3, test = "F")
``



```{r}
lin_model2 <- glm(R_moment_2 ~ Re + St + Fr, data = clean)
cv.glm(lin_model2, data = clean)$delta[1]
lin_model3 <- glm(R_moment_2 ~ Re + St + St^2 +  Fr, data = clean)
cv.glm(lin_model2, data = clean)$delta[1]
lin_model4 <- glm(R_moment_2 ~ St, data = clean)
cv.glm(lin_model4, data = clean)$delta[1]
```



```{r}
m2glm1 = glm(R_moment_2 ~ Re + Fr + St + St*Fr +  Re*St*Fr, data = clean)
cv.glm(m2glm1, data = clean)$delta[1] ### 10840
```

## 360 MSE MODEL vvv

```{r}
m2glm5 = glm(R_moment_2 ~ Re + Fr + St + St*Fr +  Re*St*Fr +Re*log(St)*Fr, data = clean)
cv.glm(m2glm5, data = clean)$delta[1] ## MSE = 359.2733

summary(m2glm5)
```

This model does not work for moments 3 or 4

```{r}
m2glm10 = glm(R_moment_2 ~ Re + Fr + St + St*Fr +  Re*St*Fr + Re*log(St)*Fr, data = clean)
cv.glm(m2glm5, data = clean)$delta[1] ## MSE = 359.2733

summary(m2glm10)
```

```{r}
m2glm7 = glm(R_moment_2 ~ Re + Fr + St + St:log(St):Fr, data = clean)
cv.glm(m2glm7, data = clean)$delta[1] 
```


```{r doesntwork}
m2glm5 = glm(R_moment_2 ~ Re + Fr + St + Re90:Fr0.052:log(St), data = clean) ##invalid
```


```{r}
full.model <- glm(R_moment_2 ~ Re + Fr + St + St*Fr + St*Re + Re*Fr + log(as.numeric(Re))*Fr + Re*St*Fr, data = clean)

step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
cv.glm(step.model, data = clean)$delta[1]
summary(step.model)
```



```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_5 ~ Re + Fr + poly(St, degree), data = clean)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```

## Moment 3

```{r}
ggplot(data = clean, aes(x= St, y = R_moment_4, color = Fr)) +
  geom_point()
ggplot(data = clean, aes(x= Fr, y = R_moment_3, color = Re)) +
  geom_point()
ggplot(data = clean, aes(x= Re, y = R_moment_3, color = St)) +
  geom_point()
```

```{r}
full.model3 <- glm(R_moment_3 ~ Re + Fr + St + St*Fr + St*Re + Re*Fr +Re*log(St)*Fr, data = clean)
cv.glm(step.model3, data = clean)$delta[1] 
```
10840.18 same with/without the interaction without the log, and with adding up to exponential 3 of each alone


```{r}
rm2_over = data.train %>%
  filter(Fr == 0.052 & Re == 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))

rm2_under = data.train %>%
  filter(Fr != 0.052 | Re != 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))
```

```{r}
rm2_under_glm = glm(R_moment_3 ~ Re + Fr + St, data = rm2_under)
cv.glm(rm2_under_glm, data = rm2_under)$delta[1]
```
```{r}
rm2_under_glm = glm(R_moment_3 ~ Re + Fr + St, data = rm2_over)
cv.glm(rm2_under_glm, data = rm2_over)$delta[1]
```

```{r}
re_over = data.train %>%
  filter(Re == 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))

re_under = data.train %>%
  filter(Re != 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))
```

```{r}
re_under_glm2 = glm(R_moment_2 ~ Fr + Re + St + St^2 + St^3  + Re*log(St)*Fr + Re*St^3*Fr, 
                   data = re_under) ##looks liek a cube but doesnt change shit
cv.glm(re_under_glm2, data = re_under)$delta[1]

```

```{r}
rm2_over %>%
  ggplot(aes(x = St, y = R_moment_2, color = Re, shape = Fr)) +
  geom_smooth(method = "lm", formula = "y ~ log(x)") +
  geom_point()

rm2_under %>%
  ggplot(aes(x = St, y = R_moment_2, color = Re, shape = Fr)) +
    geom_smooth(method = "lm", formula = "y ~ log(x)") +
    geom_point()
```


```{r}
ggplot(data = re_under, aes(St, R_moment_2, color = Fr)) + geom_point()
ggplot(data = re_over, aes(St, R_moment_2, color = Fr)) + geom_point()
ggplot(data = clean, aes(Re, R_moment_3, color = Fr)) + geom_point()
```


```{r}
re_over_glm2 = glm(R_moment_2 ~ Fr + St + St*Fr + log(St)*Fr, 
                   data = re_over)
suppressWarnings(cv.glm(re_over_glm2, data = re_over)$delta[1])
```


## Results


## Conclusion