---
title: "Predicting Turbulence Simulations"
author: "Enzo Moraes Mescall, Martin Olarte, Charlotte Coudert"
date: "2022-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Report should be no more than 4 pages
```


```{r libraries, echo = FALSE, results = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(splines)
library(gam)
library(boot)
```

```{r data}
data.train = read.csv("data-train.csv")
data.test = read.csv("data-test.csv")
```


## Introduction


## Methodology

We wanted a method to quickly perform cross validation on a candidate model and estimate its test MSE, so we created a general function with that purpose.

```{r}
cv_mse = function(input_model, k = 5, iter = 25, data = clean) {
    sum_mse = 0
    for(i in 1:iter) {
      sum_mse = sum_mse + cv.glm(input_model, data = clean, K = k)$delta[1]
    }
    avg_mse = sum_mse/iter
    return(avg_mse)
}
```


After this, we created a function that would resemble our final product and take in a tuple of St, Re, and Fr and then return the four predicted moments. We reserved the global variable names `model_mX` for this final function.

```{r}
# temp models
predictive_model = function(test_St, test_Re, test_Fr) {
  newdata = data.frame(list(St = test_St, Re = as.factor(test_Re), Fr = as.factor(test_Fr)))
  
  m1 = predict(model_m1, newdata =  newdata)
  m2 = predict(model_m2, newdata =  newdata)
  m3 = predict(model_m3, newdata =  newdata)
  m4 = predict(model_m4, newdata =  newdata)
  
  return(c(m1, m2, m3, m4))
}
```

```{r}
summary(data.train)
```

```{r}
clean = data.train %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))
```


```{r}
glm(R_moment_1~ .
```

```{r}
ggplot(data = clean, aes(x= St, y = Re, color = Fr)) +
  geom_point() + geom_jitter()
```

```{r}
ggplot(data = clean, aes(x= St)) +
  geom_histogram()
ggplot(data = clean, aes(x= Fr)) +
  geom_bar()
ggplot(data = clean, aes(x= Re)) +
  geom_bar()
```


```{r gen-linear}
lin_model <- glm(R_moment_1 ~ Re + St + Fr, data = clean)
summary(lin_model)

par(mfrow = c(2,2))
plot(lin_model)
```
## MOMENT 2


```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_1 ~ Re + Fr + poly(St, degree), data = data.train)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```


```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_1 ~ Re + Fr + poly(St, degree), data = clean)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```

```{r}
ggplot(data = clean, aes(x= St, y = R_moment_2, color = Fr)) +
  geom_point()
ggplot(data = clean, aes(x= Fr, y = R_moment_2, color = Re)) +
  geom_point()
ggplot(data = clean, aes(x= Re, y = R_moment_2, color = St)) +
  geom_point()
```


```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_2 ~ Re + Fr + poly(St, degree), data = clean)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```

``{r}
gam.m1 <- gam(R_moment_2 ~ s(St, 5) + Re+ Fr, data = clean)
preds1 <- predict(gam.m1, newdata = clean)

gam.m2 <- gam(R_moment_2 ~ ns(St, 5) + Re+ Fr, data = clean)
gam.m3 <- gam(R_moment_2 ~ s(St, 3)  + Re+ Fr, data = clean)
anova(gam.m1, gam.m2, gam.m3, test = "F")
``



```{r}
lin_model2 <- glm(R_moment_2 ~ Re + St + Fr, data = clean)
cv.glm(lin_model2, data = clean)$delta[1]
lin_model3 <- glm(R_moment_2 ~ Re + St + St^2 +  Fr, data = clean)
cv.glm(lin_model2, data = clean)$delta[1]
lin_model4 <- glm(R_moment_2 ~ St, data = clean)
cv.glm(lin_model4, data = clean)$delta[1]
```



```{r}
m2glm1 = glm(R_moment_2 ~ Re + Fr + St + St*Fr +  Re*St*Fr, data = clean)
cv.glm(m2glm1, data = clean)$delta[1] ### 10840
```

## 360 MSE MODEL vvv

```{r}
m2glm5 = glm(R_moment_2 ~ Re + Fr + St + St*Fr +  Re*St*Fr +Re*log(St)*Fr, data = clean)
cv.glm(m2glm5, data = clean)$delta[1] ## MSE = 359.2733

summary(m2glm5)
```

This model does not work for moments 3 or 4

```{r}
m2glm10 = glm(R_moment_2 ~ Re + Fr + St + St*Fr +  Re*St*Fr + Re*log(St)*Fr, data = clean)
cv.glm(m2glm5, data = clean)$delta[1] ## MSE = 359.2733

summary(m2glm10)
```

```{r}
m2glm7 = glm(R_moment_2 ~ Re + Fr + St + St:log(St):Fr, data = clean)
cv.glm(m2glm7, data = clean)$delta[1] 
```


```{r doesntwork}
m2glm5 = glm(R_moment_2 ~ Re + Fr + St + Re90:Fr0.052:log(St), data = clean) ##invalid
```


```{r}
full.model <- glm(R_moment_2 ~ Re + Fr + St + St*Fr + St*Re + Re*Fr + log(as.numeric(Re))*Fr + Re*St*Fr, data = clean)

step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
cv.glm(step.model, data = clean)$delta[1]
summary(step.model)
```



```{r}
poly_mse_cv = c()
for (degree in 1:10) {
  degree_i = glm(R_moment_5 ~ Re + Fr + poly(St, degree), data = clean)
  poly_mse_cv[degree] = cv.glm(degree_i, data = clean)$delta[1]
}
min(poly_mse_cv)
```

## Moment 3

```{r}
ggplot(data = clean, aes(x= St, y = R_moment_4, color = Fr)) +
  geom_point()
ggplot(data = clean, aes(x= Fr, y = R_moment_3, color = Re)) +
  geom_point()
ggplot(data = clean, aes(x= Re, y = R_moment_3, color = St)) +
  geom_point()
```

```{r}
full.model3 <- glm(R_moment_3 ~ Re + Fr + St + St*Fr + St*Re + Re*Fr +Re*log(St)*Fr, data = clean)
cv.glm(step.model3, data = clean)$delta[1] 
```
10840.18 same with/without the interaction without the log, and with adding up to exponential 3 of each alone


```{r}
rm2_over = data.train %>%
  filter(Fr == 0.052 & Re == 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))

rm2_under = data.train %>%
  filter(Fr != 0.052 | Re != 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))
```

```{r}
rm2_under_glm = glm(R_moment_3 ~ Re + Fr + St, data = rm2_under)
cv.glm(rm2_under_glm, data = rm2_under)$delta[1]
```
```{r}
rm2_under_glm = glm(R_moment_3 ~ Re + Fr + St, data = rm2_over)
cv.glm(rm2_under_glm, data = rm2_over)$delta[1]
```

```{r}
re_over = data.train %>%
  filter(Re == 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))

re_under = data.train %>%
  filter(Re != 90) %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re))
```

```{r}
re_glm2 = glm(R_moment_3 ~ Fr + Re + St + St^2 + St^3 + St^4  + Re*log(St)*Fr, 
                   data = clean) ##looks like a cube but doesnt change shit
suppressWarnings(cv.glm(re_glm2, data = clean)$delta[1])

re_glm2 = glm(R_moment_3 ~ Fr + Re  + Re*log(St)*Fr, 
                   data = clean) 
suppressWarnings(cv.glm(re_glm2, data = clean)$delta[1])

re_glm2 = glm(R_moment_3 ~ Fr + Re  + Re*log(St)*Fr, 
                   data = clean)
suppressWarnings(cv.glm(re_glm2, data = clean)$delta[1])


```



```{r}

for (i in 2:10) {
  lm = glm(R_moment_3 ~ bs(St, df = i, degree = 2)*Re*Fr + Fr + Re + St + St^2 + St^3 + St^4  + Re*log(St)*Fr, data = clean)
  errors[i] = suppressWarnings(cv.glm(clean, lm)$delta[1])
}
plot(2:10, errors[-1], type="b", xlab="df", ylab="CV error", main = 'CV Error vs Degrees of Freedom with Degree=2 Splines')
abline(v = which.min(errors), col = "red")

spline_with_truple = glm(R_moment_3 ~ bs(St, df = 1, degree = 2)*Re*Fr + Fr + Re + St + St^2 + St^3 + St^4  + Re*log(St)*Fr + Re*St^3*Fr, data = clean)
cv.glm(spline_with_truple, data = clean)$delta[1]

##unhelpful

```

```{r}
rm2_over %>%
  ggplot(aes(x = St, y = R_moment_2, color = Re, shape = Fr)) +
  geom_smooth(method = "lm", formula = "y ~ log(x)") +
  geom_point()

rm2_under %>%
  ggplot(aes(x = St, y = R_moment_2, color = Re, shape = Fr)) +
    geom_smooth(method = "lm", formula = "y ~ log(x)") +
    geom_point()
```


```{r}
ggplot(data = re_under, aes(St, R_moment_3, color = Fr)) + labs(title = "2nd Moment Re = 90") + geom_point()
ggplot(data = re_over, aes(St, R_moment_2, color = Fr)) +  labs(title = "2nd Moment Re > 90") + geom_point()
ggplot(data = clean, aes(Re, R_moment_3, color = Fr)) + labs(title = "3rd Moment Fr, Re Distribution") +geom_point()
```


```{r}
re_over_glm2 = glm(R_moment_2 ~ Fr + St + St*Fr + log(St)*Fr, 
                   data = re_over)
suppressWarnings(cv.glm(re_over_glm2, data = re_over)$delta[1])
```
```{r}
model_m1 = glm(R_moment_1 ~ log(St)*Fr*Re + St*Fr*Re, data = clean)
model_m2 = glm(R_moment_2 ~ log(St)*Fr*Re + St*Fr*Re, data = clean)
```

```{r}
model_m1
summ(model_m2)
```

```{r}
test = data.train %>%
  mutate(Fr = as.factor(Fr),
         Re = as.factor(Re),
         St = log(St))
test_model = glm(R_moment_2 ~ St+ Fr+ Re + St*Fr*Re, 
                   data = test)
summary(test_model)
cv.glm(test_model, data = test)$delta[1]
```

```{r}
ggplot(data = clean
       , aes(x = R_moment_3, y = R_moment_4, color = Fr)) + geom_point()
```


## Introduction

An introduction outlining key research objectives and how your model achieves such objectives.

In this project we explore the distribution and clustering of particles in an idealized turbulence as effected by Reynolds number (Re), Froude number (Fr), and Stokes number (St), which quantify the fluid turbulence, gravitational acceleration, and particle characteristics, respectively. Using data that simulated Re, Fr, and St and measured the produced clusters, quantifying them by their first 4 moments. We were able to find generalized linear model incorporating the interaction of factored Fr and St values and log(Re). This model gives us only moderate predicative accuracy but is very interpretable. 


